---
title: "EGA Dimensions"
format: html
---

# **Exploratory Graph Analysis to perform item clustering**

For the scope of this Thesis, it was decided to use EGA to select the dimensions for the Confirmatory factor analysis.

```{r}
# Tidyverse
getwd()
library(dplyr)  # Data manipulation
library(purrr)  # Functional programming
library(readr)  # Data import

# EGA
source("../DataPreparation/functions.R")
library(EGAnet)
library(sna)
```

```{r}
df <-ImportReversedDF()

# Get the names of the columns where max value is either 2 or above 6
variables_to_remove <- names(df)[sapply(df, function(x) max(x, na.rm = TRUE) %in% 2 | max(x, na.rm = TRUE) > 6)]

df <- df %>% select(-variables_to_remove,-ID, -SurveyCombination_M1, -SurveyCombination_M2, -Country,-private_sector,-empl_contract,-eng_energy,-eng_enthusiastic,-eng_timeflies)

df <- df %>% select(-telework,-losejob,-undesirable_change,-predict_earnings)
```
## Removing items with low stability

We remove items that are not found reliably in the same clusters from the dataset

```{r}
preliminarybootdata <- bootEGA(
                    df,
                    iter= 1000, 
                    seed = 42,
                    algorithm = 'louvain',
                    type = "resampling",
                    progress = FALSE,
                    )
```
```{r}
preliminarybootdata$frequency
```


```{r}
preliminaryIS <- itemStability(preliminarybootdata)
```

```{r}
names_to_remove <- names(preliminaryIS$item.stability$empirical.dimensions)[preliminaryIS$item.stability$empirical.dimensions < 0.8]

df <- df %>%
  select(-all_of(names_to_remove))
```

## EGA clustering

We perform the clustering with stable variables and then save the results.

```{r}
bootdata <- bootEGA(
                    df,
                    iter= 1000, 
                    seed = 42,
                    algorithm = 'louvain',
                    type = "resampling",
                    progress = FALSE,
                    )
```
```{r}
bootdata$frequency
```


```{r}
IS <- itemStability(bootdata)
```

```{r}
dfitemscluster <- tibble(
  names = names(IS$membership$empirical),
  values = IS$membership$empirical
)

# Import the CSV file
used_variables <- ImportUsedvar()  %>%
  select(variable, question , dimension)

# Perform a join operation
dfitemscluster <- dfitemscluster %>%
  left_join(used_variables, by = c("names" = "variable"))
```

```{r}
write_csv(dfitemscluster, "../../data/processed/EGA_itemclusters.csv")
```

After saving the data, an EGAdimensions.csv was made manually to compare the EGA dimensions with EWCS ones and to assign a proper name to these dimensions.